---
title: "iteration_list"
author: "Lucia Wang"
date: "2023-10-31"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
library(purrr)
set.seed(12345)
```

## lists
lists can be useful to store data, and for loops can be used to write over them - use functions.

lists can contain anything! even a list. -> essentially a data frame?

main thing is it has an index 

Recall: 
```{r}
vec_numeric = 1:4
vec_char = c("My", "name", "is", "Jeff")

tibble(num = vec_numeric,
       char = vec_char)
```

what if you add  stuff with different lengths - cant put df together, but you can make a list
```{r}
l = list(
  vec_numeric = 1:5,
  vec_char = LETTERS,
  matrix = matrix(1:10, nrow=5, ncol=2),
  summary = summary(rnorm(100))
)
```

accessing lists with $ or [[]]
```{r}
l$vec_char 
l[[2]]
l[["summary"]]
```

## for loops
```{r}
list_norm_samples = list(
  a = rnorm(20,1,5),
  b= rnorm(20, 0, 7),
  c= rnorm(20, 20, 1),
  d = rnorm(20, -45, 13)
)

# recall mean and sd fn from last class...
mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) == 1) {
    stop("Cannot be computed for length 1 vectors")
  }
  
  mean_x = mean(x)
  sd_x = sd(x)

  tibble(
    mean = mean_x, 
    sd = sd_x
  )
}
```

you could do it individually.. but its annoying
```{r}
mean_and_sd(list_norm_samples$a)
mean_and_sd(list_norm_samples$b)

```

loop over the list instead
```{r}

output = vector("list", length = 4)

for (i in 1:4) {
  output[[i]] = mean_and_sd(list_norm_samples[[i]])
}
```
## map

`map()` does the same thing as for loop with basic structure 
`output = map(input, f)`. easier to read but same thing

```{r}
output_mean_sd=map(list_norm_samples, mean_and_sd)
output_summary = map(list_norm_samples, summary)
```

## list columns

```{r}
listcol_df = 
  tibble(
    name = c("a", "b", "c", "d"),
    samp = list_norm_samples
  )

# this is a lot to type 4 times
mean_and_sd(listcol_df$samp[[1]])
mean_and_sd(listcol_df$samp[[2]])

# do this instead
map(listcol_df$samp, mean_and_sd)

# even update the column
listcol_df |>
  mutate(mean_sd = map(samp, mean_and_sd),
         median = map(samp, median)
         ) |>
  select(name, mean_sd) |>
  unnest(mean_sd)
```

revisiting nsduh data.
this function takes in the html and table num to get you a table
```{r}
nsduh_table <- function(html, table_num, table_name) {
  
  table = 
    html |> 
    html_table() |> 
    nth(table_num) |>
    slice(-1) |> 
    select(-contains("P Value")) |>
    pivot_longer(
      -State,
      names_to = "age_year", 
      values_to = "percent") |>
    separate(age_year, into = c("age", "year"), sep = "\\(") |>
    mutate(
      year = str_replace(year, "\\)", ""),
      percent = str_replace(percent, "[a-c]$", ""),
      percent = as.numeric(percent),
      name = table_name) |>
    filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
  
  table
}

```

```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

# if you want to withdraw only one
nsduh_table(nsduh_html, 1, "marj")

# or use loop

table_in = list(1,4,5)
name_in = list("marj", "cocaine", "heroin")

output = vector("list", 3)

for (i in 1:3) {
  output[[i]] = nsduh_table(nsduh_html, table_in[[i]], name_in[[i]])
}

output
```

try again.. with maps
```{r}
nsduh_df =
  tibble(
    name = c("marj", "cocaine", "heroin"),
    number = c(1,4,5)
  )

# simplify your import function
nsduh_table2 <- function(html, table_num) {
  
  table = 
    html |> 
    html_table() |> 
    nth(table_num) |>
    slice(-1) |> 
    select(-contains("P Value")) |>
    pivot_longer(
      -State,
      names_to = "age_year", 
      values_to = "percent") |>
    separate(age_year, into = c("age", "year"), sep = "\\(") |>
    mutate(
      year = str_replace(year, "\\)", ""),
      percent = str_replace(percent, "[a-c]$", ""),
      percent = as.numeric(percent)) |>
    filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
  
  table
}

# applying param 2 with param 3 on each param 1
map(nsduh_df$number, nsduh_table2, html=nsduh_html)

nsduh_df =
  tibble(
    name = c("marj", "cocaine", "heroin"),
    number = c(1,4,5)
  ) |>
  mutate(
    table = map(number, nsduh_table2, html=nsduh_html)
  ) |>
  unnest(table)

```
not necessarily any faster but its another method

